### 2016-11-09 datamining/summery.tex
added definitions of baysian

\subsection{Decomosable graphical models}
Decomposable models are graphical models that have explicit formulas for the
maximum likelihood estimates.
This is a convenient property from a computational viewpoint.
If we only have to fit one model this is perhaps not so important,
but when we have little prior knowledge we typically have to search a
potentially large space of possible models.
Decomposable models are very easy to characterize by their independence graphs.
They have triangulated independence graphs: their independence graphs have no
chordless cycles of length greater than three. A cycle is called chordless
if no other than successive pairs of vertices in the cycle are adjacent. 
\subsection{Deviance and likelhood ratio test}
The deviance of a fitted model compares the log-likelihood of the fitted model
to the log-likelihood of the saturated model.
The larger the model deviance, the poorer the fit.
The likelihood of a model $M$ is
$L(\overset{\wedge}{P}(x);n(x))=\underset{x}{\prod}\overset{\wedge}{P}(x)^{n(x)}$
and the log likelihood of a model $M$ is
$\underset{x}{\sum} n(x)\log \overset{\wedge}{P}(x)$
$L^i=L(\overset{\wedge}{P}^{M_i})$ is the value of the log likelihood function
valutated at $\overset{\wedge}{P}^{M_i}$; the ML estimates of $P$ under $M_i$.
Let $M_0 \subseteq M_1$; ie $M_0$ can be obtained from $M_1$ by setting
additional $u$-terms to zero (more restrictions). The deviance between $M_0$ and
$M_1$ is $dev(M_0) - dev(M_1) = 2(L^1 - L^0)$.
For large $N$ $2(L^1-L^0) \approx M_0 \chi^2_v$ where the degrees of freedom $v$
is equal to the number of additional restrictions of $M_0$. 
We reject the null hypothesis that $M_0$ is the true model when:
$2(L^1-L^0)>\chi^2_{v;\alpha}$. 
\section{Baysian networks}

\section{Freq item set mining}
\end{multicols}
\end{document}

>>>>

\subsection{Decomosable graphical models}
Decomposable models are graphical models that have explicit formulas for the
maximum likelihood estimates.
This is a convenient property from a computational viewpoint.
If we only have to fit one model this is perhaps not so important,
but when we have little prior knowledge we typically have to search a
potentially large space of possible models.
Decomposable models are very easy to characterize by their independence graphs.
They have triangulated independence graphs: their independence graphs have no
chordless cycles of length greater than three. A cycle is called chordless
if no other than successive pairs of vertices in the cycle are adjacent. 
\subsection{Deviance and likelhood ratio test}
The deviance of a fitted model compares the log-likelihood of the fitted model
to the log-likelihood of the saturated model.
The larger the model deviance, the poorer the fit.
The likelihood of a model $M$ is
$L(\overset{\wedge}{P}(x);n(x))=\underset{x}{\prod}\overset{\wedge}{P}(x)^{n(x)}$
and the log likelihood of a model $M$ is
$\underset{x}{\sum} n(x)\log \overset{\wedge}{P}(x)$
$L^i=L(\overset{\wedge}{P}^{M_i})$ is the value of the log likelihood function
valutated at $\overset{\wedge}{P}^{M_i}$; the ML estimates of $P$ under $M_i$.
Let $M_0 \subseteq M_1$; ie $M_0$ can be obtained from $M_1$ by setting
additional $u$-terms to zero (more restrictions). The deviance between $M_0$ and
$M_1$ is $dev(M_0) - dev(M_1) = 2(L^1 - L^0)$.
For large $N$ $2(L^1-L^0) \approx M_0 \chi^2_v$ where the degrees of freedom $v$
is equal to the number of additional restrictions of $M_0$. 
We reject the null hypothesis that $M_0$ is the true model when:
$2(L^1-L^0)>\chi^2_{v;\alpha}$. 
\section{Baysian networks}
\subsection{Definition and notation}
A directed graph is a pair $G = (K, E)$, where $K$ is a set of vertices and $E$
is a set of edges with ordered pairs of vertices.
If there is an arrow from $i$ to $j$,
then we write this as $i \to j$, or equivalently as $(ij) \in E$.
We restrict attention to directed graphs with no directed cycles,
i.e. acyclic directed graphs (DAGs). If $i \to j$,
then $X_i$ is called a parent of $X_j$ ,
and $X_j$ is called a child of $X_i$ .
The set of coordinates of the parents of $X_j$ is denoted $pa(j)$, so
$X_{pa(j)}$ denotes the set of parents of $X_j$ .
If there is a directed path from $i$ to $j$,
then $X_i$ is called an ancestor of $X_j$ .
The set of coordinates of the ancestors of $X_j$ is denoted an $(j)$.
The absence of any directed cycles is equivalent to the existence of an
ordering of the nodes ${1, 2, \dots , k}$ such that $i \to j$ only when $i < j$.
In other words, there exists an ordering of the nodes such that arrows point
only from lower-numbered nodes to higher-numbered nodes.
Suppose that a priori knowledge tells us the variables can be labeled
$X_1 , X_2 , \dots , X_k$ such that $X_i$ is prior to $X_{i+1}$.
Corresponding to this ordering we can factorize the joint density of
$X_1 , X_2 , \dots , X_k$ as $P(X)=P(X_1)P(X_2|X_1)\dots P(X_k|X_{k-1},X_{k-2},\dots,X_1)$
In contructing a DAG, an arrow is drawn from $i$ to $j$, where $i<j$, unless
$P(X_j|X_{j-1}\dots,X_1)$ does not depend on $X_i$.
This is the key difference between DAGs and undirected graphs.
In both types of graph a missing edge between $X_i$ and $X_j$ is equivalent to
a conditional independence relation between $X_i$ and $X_j$.
In undirected graphs they are conditionally independent given all the remaining
variables, whereas in DAGs they are conditionally independent given all prior
variables. 
We can write the joint density:
$P(X)=\underset{i=1}{\overset{k}{\prod}}P(X_i|X_{pa(i)})$.
and the conditional indepencence as: $i \indep j|an({i,j})$
\end{multicols}
\end{document}
